DALEX_TechQualityControl

### 2021-03-11 DALEX_tech_QAQC.Rmd
multiple thresholds rmd initial commit

---
title: "DALEX_TechQualityControl"
author: "Scott Coffin"
date: "2/23/2021"
output: 
  html_document:
    code_folding: hide
    theme: sandstone
    toc: yes
    toc_float: yes
    toc_depth: 6
    number_sections: true
  word_document:
    toc: yes
---
```{r global-options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      warning=FALSE, message=FALSE, echo = FALSE)
```

```{r message=FALSE, warning=FALSE}
# load required packages
library(rsample)
library(dplyr)
library(DALEX)
library(tidyverse)
library(tigerstats)
library(caret)
library(skimr)
library(ggeffects)
library(knitr)
library(doParallel)
library(modelplotr)
library(gridExtra)
```
# Setup
## Data Import
```{r}
# Load finalized dataset.
aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000) #%>% 
  #mutate(ID = paste0("ID",row_number()))

@@

# final cleanup and factoring  
aoc_z$Species <- as.factor(paste(aoc_z$genus,aoc_z$species)) #must make value 'Species" (uppercase)
aoc_z$Group <- as.factor(aoc_z$organism.group) #must make value "Group"
aoc_z$Group <- fct_explicit_na(aoc_z$Group) #makes sure that species get counted even if they're missing a group
```
## Data Selection
```{r}
df <- aoc_z %>% 
  filter(tech.tier.zero == "N") %>% #gives studies that pass technical quality criteria
  filter(risk.tier.zero == "N") %>%  #only studies applicable for risk assessment. VERY RESTRICTIVE 
  dplyr::select(c(organism.group, exposure.duration.d, lvl1_f, dose.um3.mL.master, dose.mg.L.master, dose.particles.mL.master, life.stage, bio.org, polymer, shape, size.length.um.used.for.conversions, treatments, effect, effect.metric, particle.volume.um3, density.mg.um.3, environment, exposure.route, lvl2_f)) %>% 
  mutate(effect.metric = as.character(effect.metric)) %>% 
  mutate(effect.metric = (case_when(
    effect.metric == "NONEC" ~ "NOEC",
    effect.metric == "HONEC" ~ "NOEC",
    effect.metric == "LOEC" ~ "LOEC",
    effect.metric == "LC50" ~ "LC50",
    effect.metric == "EC50" ~"EC50",
    effect.metric == "EC10" ~ "EC10"
  ))) %>% 
  mutate(effect.metric = replace_na(effect.metric,"not_available")) %>% 
  mutate(effect.metric = as.factor(effect.metric)) %>% 
  drop_na() %>%  #drop missing
mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
  mutate(effect_10 = case_when( #convert ordinal to numeric
      effect == "Y" ~ 1,
      effect == "N" ~ 0
    )) %>%
  mutate(effect_10 = factor(effect_10)) %>% 
  dplyr::select(-(effect))

@@

mp_classif_final_model <- model_performance(explainer_final_model)
mp_classif_final_model
```
#### Performance
Performance metrics for the final model are below.

@@

# Machine Settings

>>>>

---
title: "DALEX_TechQualityControl"
author: "Scott Coffin"
date: "2/23/2021"
output: 
  html_document:
    code_folding: hide
    theme: journal
    toc: yes
    toc_float: yes
    toc_depth: 6
    number_sections: true
    includes:
      after_body: footer.html
  word_document:
    toc: yes
---
```{r global-options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      warning=FALSE, message=FALSE,time_it = TRUE) #report time to knit for all chunks

#knit time reporter
all_times <- list()  # store the time for all chunks in a list
knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      res <- difftime(Sys.time(), now)
      all_times[[options$label]] <<- res
    }
  }
}))
```

```{r message=FALSE, warning=FALSE, include=FALSE}
# load required packages
library(rsample)
library(dplyr)
library(DALEX)
library(tidyverse)
library(tigerstats)
library(caret)
library(skimr)
library(ggeffects)
library(knitr)
library(doParallel)
library(modelplotr)
library(gridExtra)
```

```{r include=FALSE}
# Load finalized dataset.
aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000) #%>% 
  #mutate(ID = paste0("ID",row_number()))

@@

# final cleanup and factoring  
aoc_z$Species <- as.factor(paste(aoc_z$genus,aoc_z$species)) #must make value 'Species" (uppercase)
aoc_z$Group <- as.factor(aoc_z$organism.group) #must make value "Group"
aoc_z$Group <- fct_explicit_na(aoc_z$Group) #makes sure that species get counted even if they're missing a group

# save data object
#saveRDS(aoc_z, file = "aoc_z.Rda")
```
# Data Selection

**Important**: this dataset has ONLY been filtered for 'red' criteria for technical quality.

```{r}
df <- aoc_z %>% 
  filter(tech.tier.zero == "N") %>% #gives studies that pass technical quality criteria
  filter(risk.tier.zero == "N") %>%  #only studies applicable for risk assessment. VERY RESTRICTIVE 
  dplyr::select(c(organism.group, exposure.duration.d, lvl1_f, dose.um3.mL.master, dose.mg.L.master, dose.particles.mL.master, life.stage, bio.org, polymer, shape, size.length.um.used.for.conversions, treatments, effect, effect.metric, particle.volume.um3, density.mg.um.3, environment, exposure.route, lvl2_f)) %>% 
  mutate(effect.metric = as.character(effect.metric)) %>% 
  mutate(effect.metric = (case_when(
    effect.metric == "NONEC" ~ "NOEC",
    effect.metric == "HONEC" ~ "NOEC",
    effect.metric == "LOEC" ~ "LOEC",
    effect.metric == "LC50" ~ "LC50",
    effect.metric == "EC50" ~"EC50",
    effect.metric == "EC10" ~ "EC10"
  ))) %>% 
  mutate(effect.metric = replace_na(effect.metric,"not_available")) %>% 
  mutate(effect.metric = as.factor(effect.metric)) %>% 
  drop_na() %>%  #drop missing
mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
  mutate(effect_10 = case_when( #convert ordinal to numeric
      effect == "Y" ~ 1,
      effect == "N" ~ 0
    )) %>%
  mutate(effect_10 = factor(effect_10)) %>% 
  dplyr::select(-(effect))

@@

mp_classif_final_model <- model_performance(explainer_final_model)
mp_classif_final_model
```

```{r final model save}
#save model
saveRDS(final_model, file="final_rf_model.rds")
#load model
# final_model <- readRDS("final_rf_model.rds")
```

#### Performance
Performance metrics for the final model are below.

@@

##### What parameters bias the model?
Test residuals for size, dose, etc.

# Machine Settings
Time to knit:
```{r}
all_times
```
